---
title: "Obtaining and cleaning geographic data"
output: 
    rmarkdown::html_vignette
bibliography: library.bib
vignette: >
  %\VignetteIndexEntry{gdata_cleaning}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 4,
  fig.height = 4
)
library(dplyr)
options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r setup}
library(ntbox)
library(dplyr)
```



## Geographic data

In this example we will dowonland ocrrence data for *Leopardus wiedii* 
from the Global Biodiversity Information Facility (https://www.gbif.org) 
and explore what is the provenance and date of collecting these points. Then
we will remove sptial duplicates.


## Searching the data

```{r gbif, echo=TRUE,eval=T}
dAll_a <- ntbox::searh_gbif_data(genus = "Leopardus",
                                 species = "wiedii",
                                 occlim = 5000,
                                 leafletplot = TRUE)
```
```{r}
cat("Total number of occurrence data found:",nrow(dAll_a))
```

## Data curation 

In this example we filter those records starting in 1950.

```{r echo=TRUE, eval=TRUE}
dAll_b <- ntbox::clean_dup(dAll_a,longitude = "longitude","latitude",
                           threshold = 0)
cat("Total number of occurrence data affter cleanining spatial duplicates:",
    nrow(dAll_b))

dAll_c <- dAll_b %>% dplyr::filter(year>=1950)
cat("Total number of occurrence data for periods >=1950:",
    nrow(dAll_c))

```


```{r }
m <- leaflet::leaflet(dAll_c)
m <- m %>% leaflet::addTiles()
m <- m %>% leaflet::addCircleMarkers(lng = ~longitude, 
                                     lat = ~latitude, 
                                     popup = ~leaflet_info, 
                                     fillOpacity = 0.25, 
                                     radius = 7,col="green")
m
```


Remove wired occurrences. Click on the pop-up to display gbif information (available when the downloaded data comes from `search_gbif` function), the points that are outside the distribution are the one on San Francisco (this comes from a collection; rowID=632), the record on Florida (rowID=508,489), and the ones that in the sea (544,595). 


```{r echo=TRUE,eval=TRUE,fig.width=4,fig.height=4}
# Indixes of the wired data (can change depending the date of the gbif query)
to_rmIDs <- c(636,493,512,544,595)
to_rm <- which(dAll_c$ntboxID %in% to_rmIDs)
dAll <- dAll_c[-to_rm,]
m <- leaflet::leaflet(dAll)
m <- m %>% leaflet::addTiles()
m <- m %>% leaflet::addCircleMarkers(lng = ~longitude, 
                                     lat = ~latitude, 
                                     popup = ~leaflet_info, 
                                     fillOpacity = 0.25, 
                                     radius = 7,col="green")
m
```


